\documentclass[a4paper]{article}

\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\usepackage{minted}
\usepackage{multirow}
\usepackage{float}
\usepackage{todonotes}
\usepackage{forest}
\usepackage{xstring}
\usepackage{xcolor}
\usepackage[most]{tcolorbox}
\usepackage{xparse}
\usepackage{natbib}
\usepackage{tikz}

\usetikzlibrary{graphs,graphdrawing}
\usegdlibrary{layered}

% External Graphviz support
\input{/home/donkarlo/Dropbox/repo/nd_ai_project/src/nd_ai/knoweldge_representation/language/formal/computer/programming/tex/latex/code_clips/external_graphviz.tex}

% Semantic tag macros
\input{/home/donkarlo/Dropbox/repo/nd_ai_project/src/nd_ai/knoweldge_representation/language/formal/computer/programming/tex/latex/parts/control_sequence/kind/semtag/semtag.tex}

% Link macros
\input{/home/donkarlo/Dropbox/repo/nd_ai_project/src/nd_ai/knoweldge_representation/language/formal/computer/programming/tex/latex/code_clips/seehere.tex}

\title{Waveform}
\author{}
\date{}

\begin{document}
    \maketitle
    
    
    \section{FitzHugh--Nagumo model}
        
        The FitzHugh--Nagumo model (FHN) describes a prototype of an excitable system, such as a neuron.
        It is a simplified dynamical model of neuronal excitability and action-potential-like behavior.
        
        \begin{equation}
            \dot{v}
            =
            c
            \Bigl(
            v
            -
            \frac{v^{3}}{3}
            +
            w
            +
            I
            \Bigr)
        \end{equation}
        
        \begin{equation}
            \dot{w}
            =
            -
            \frac{1}{c}
            \Bigl(
            v
            -
            a
            +
            bw
            \Bigr)
        \end{equation}
        
        \cite{cebrian-lacasa-2025-six-decades-of-the-fitzhugh-nagumo-model-a-guide-through-its-spatio-temporal-dynamics-and-influence-across-disciplines}
        reviews six decades of research on the FitzHugh--Nagumo model as a simplified but powerful description of excitable dynamics.
        It shows that the model captures not only neuronal excitability and oscillations, but also bistability, traveling waves, synchronization, and spatial pattern formation.
        The paper emphasizes that the model has become useful beyond neuroscience, including cardiac dynamics, biology, electronics, optics, mathematics, and physics.
        Its main contribution is to organize the known dynamical regimes and bifurcations of the model, especially in spatially extended and coupled systems.
    
    
    \section{Frequency Modulated M\"obius model}
        
        \cite{rueda-2021-a-novel-wave-decomposition-for-oscillatory-signals}
        and
        \cite{rodriguez-collado-2021-identification-of-action-potential-waveforms}
        propose a wave decomposition method for oscillatory signals based on Frequency Modulated M\"obius (FMM) waves.
        The method represents complex oscillatory patterns as a sum of interpretable wave components with parameters controlling amplitude, phase, location, and shape.
        For action-potential-like curves, its main value is that it provides a citable mathematical basis for approximating signal waveforms by combining several FMM components.
        
        \subsection{Single FMM wave}
            
            A single FMM wave can be written as follows:
            
            \begin{equation}
                W(t)
                =
                A
                \cos
                \Biggl[
                    \beta
                    +
                    2
                    \arctan
                    \Biggl(
                    \omega
                    \tan
                    \Bigl(
                    \frac{t-\alpha}{2}
                    \Bigr)
                    \Biggr)
                    \Biggr]
            \end{equation}
            
            \begin{itemize}
                \item $W(t)$ is the value of the FMM wave at time $t$.
                
                \item $t$ is time.
                
                \item $A$ is the amplitude parameter. It controls the vertical scale of the waveform.
                
                \item $\alpha$ is the location parameter. It controls the horizontal position of the waveform on the time axis.
                
                \item $\beta$ is the phase parameter. It controls the phase shift of the waveform.
                
                \item $\omega$ is the shape parameter. It controls the sharpness, skewness, and temporal deformation of the waveform.
                
                \item $\cos(\cdot)$ is the cosine function used to generate the oscillatory waveform.
                
                \item $\arctan(\cdot)$ is the inverse tangent function used inside the phase transformation.
                
                \item $\tan(\cdot)$ is the tangent function used to transform the time variable before applying the inverse tangent.
                
                \item $\beta + 2 \arctan\bigl(\omega \tan\bigl(\frac{t-\alpha}{2}\bigr)\bigr)$ is the phase function of the FMM wave.
            \end{itemize}
        
        \subsection{Multi-component FMM waveform}
            
            A complex waveform can be modeled as the sum of several FMM waves.
            The general $m$-component model is:
            
            \begin{equation}
                V(t)
                =
                M
                +
                \sum_{j=1}^{m}
                W_j(t)
            \end{equation}
            
            Equivalently:
            
            \begin{equation}
                V(t)
                =
                M
                +
                \sum_{j=1}^{m}
                A_j
                \cos
                \Biggl[
                    \beta_j
                    +
                    2
                    \arctan
                    \Biggl(
                    \omega_j
                    \tan
                    \Bigl(
                    \frac{t-\alpha_j}{2}
                    \Bigr)
                    \Biggr)
                    \Biggr]
            \end{equation}
            
            Here, $M$ is the baseline or vertical offset of the full signal.
            In the case of an action potential, $V(t)$ represents the membrane voltage at time $t$.
        
        \subsection{Two-component FMM waveform}
            
            For two FMM components, the waveform becomes:
            
            \begin{equation}
                V(t)
                =
                M
                +
                W_1(t)
                +
                W_2(t)
            \end{equation}
            
            Equivalently:
            
            \begin{equation}
                \begin{aligned}
                    V(t)
                    =
                    M
                    &+
                    A_1
                    \cos
                    \Biggl[
                        \beta_1
                        +
                        2
                        \arctan
                        \Biggl(
                        \omega_1
                        \tan
                        \Bigl(
                        \frac{t-\alpha_1}{2}
                        \Bigr)
                        \Biggr)
                        \Biggr]
                    \\
                    &+
                    A_2
                    \cos
                    \Biggl[
                        \beta_2
                        +
                        2
                        \arctan
                        \Biggl(
                        \omega_2
                        \tan
                        \Bigl(
                        \frac{t-\alpha_2}{2}
                        \Bigr)
                        \Biggr)
                        \Biggr]
                \end{aligned}
            \end{equation}
            
            This two-component form is useful when an action-potential-like waveform is approximated by two combined FMM waves.
    
    
    \section{Construction of a single five-parameter FMM wave from two FMM waves}
        
        Assume that two FMM components are given as follows:
        
        \begin{equation}
            W_1(t)
            =
            A_1
            \cos\left[
                    \beta_1
                +
                2
                    \arctan\left(
                               \omega_1
                               \tan\left(
                                       \frac{t-\alpha_1}{2}
                    \right)
                \right)
            \right]
        \end{equation}
        
        \begin{equation}
            W_2(t)
            =
            A_2
            \cos\left[
                    \beta_2
                +
                2
                    \arctan\left(
                               \omega_2
                               \tan\left(
                                       \frac{t-\alpha_2}{2}
                    \right)
                \right)
            \right]
        \end{equation}
        
        The target waveform is obtained by adding the two components together with a baseline term:
        
        \begin{equation}
            S(t)
            =
            M_0
            +
            W_1(t)
            +
            W_2(t)
        \end{equation}
        
        The goal is to approximate this two-component waveform by a single five-parameter FMM wave:
        
        \begin{equation}
            \widehat{S}(t)
            =
            M^{\star}
            +
            A^{\star}
            \cos\left[
                    \beta^{\star}
                    +
                    2
                    \arctan\left(
                               \omega^{\star}
                               \tan\left(
                                       \frac{t-\alpha^{\star}}{2}
                    \right)
                \right)
            \right]
        \end{equation}
        
        The five parameters of the new single wave are therefore:
        
        \begin{equation}
            \theta^{\star}
            =
            \left(
                M^{\star},
            A^{\star},
                \alpha^{\star},
                \beta^{\star},
                \omega^{\star}
            \right)
        \end{equation}
        
        These parameters are estimated by minimizing the squared difference between the two-component waveform and the single five-parameter waveform:
        
        \begin{equation}
            \theta^{\star}
            =
            \arg\min_{
                M,
                A,
                \alpha,
                \beta,
                \omega
            }
            \sum_{i=1}^{N}
            \left[
                S(t_i)
                -
                \left(
                    M
                    +
                    A
                    \cos\left[
                            \beta
                            +
                            2
                            \arctan\left(
                                       \omega
                                       \tan\left(
                                               \frac{t_i-\alpha}{2}
                        \right)
                    \right)
                \right]
                \right)
            \right]^2
        \end{equation}
        
        The optimization is performed under the following parameter constraints:
        
        \begin{equation}
            A \geq 0,
            \qquad
            0 \leq \alpha < 2\pi,
            \qquad
            0 \leq \beta < 2\pi,
            \qquad
            0 < \omega \leq 1
        \end{equation}
        
        In this construction, the two original FMM waves are first summed to form the target waveform \(S(t)\). Then, instead of keeping the two separate components, a single FMM wave \(\widehat{S}(t)\) is fitted to this target waveform. The resulting wave has only five parameters: baseline \(M^{\star}\), amplitude \(A^{\star}\), location \(\alpha^{\star}\), phase \(\beta^{\star}\), and shape parameter \(\omega^{\star}\). This does not give an exact algebraic replacement for the two waves in general; it gives the best single-wave approximation according to the least-squares error.
        
        The two FMM components are first added to construct a target waveform. A single five-parameter FMM wave is then fitted to this target waveform by minimizing the sum of squared errors between the target signal and the fitted signal.
        
        \subsection{Optimization method}
            
            The parameter estimation problem was formulated as a nonlinear least-squares fitting problem. The target waveform was first constructed as the sum of two FMM components:
            
            \begin{equation}
                S(t_i)
                =
                M_0
                +
                W_1(t_i)
                +
                W_2(t_i).
            \end{equation}
            
            A single five-parameter FMM waveform was then fitted to this target signal:
            
            \begin{equation}
                \widehat{S}(t_i;\theta)
                =
                M
                +
                A
                \cos\left[
                        \beta
                        +
                        2
                        \arctan\left(
                                   \omega
                                   \tan\left(
                                           \frac{t_i-\alpha}{2}
                        \right)
                    \right)
                \right],
            \end{equation}
            
            where
            
            \begin{equation}
                \theta
                =
                \left(
                    M,
                    A,
                    \alpha,
                    \beta,
                    \omega
                \right).
            \end{equation}
            
            The residual at each sampled time point was defined as
            
            \begin{equation}
                r_i(\theta)
                =
                S(t_i)
                -
                \widehat{S}(t_i;\theta).
            \end{equation}
            
            The optimal parameter vector was obtained by minimizing the sum of squared residuals:
            
            \begin{equation}
                \theta^{\star}
                =
                \arg\min_{\theta}
                \sum_{i=1}^{N}
                r_i(\theta)^2.
            \end{equation}
            
            In the implementation, this minimization was performed using a \textbf{hybrid global-local optimization} strategy. Differential Evolution was first used as a \textbf{derivative-free global optimizer} to search the parameter space and obtain a suitable initial estimate. This stage reduces the risk of convergence to a poor local minimum. The resulting estimate was then refined using bounded nonlinear least-squares optimization. Therefore, the final five-parameter FMM wave is obtained by fitting a single FMM component to the two-component target waveform through least-squares error minimization.
            
            \bibliographystyle{plainnat}
            \bibliography{/home/donkarlo/Dropbox/repo/nd_shared_research/bibliography/bibliography.bib}
\end{document}