\documentclass[a4paper]{article}
\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\usepackage{hyperref}

\title{Autobiographical}
\begin{document}
    \maketitle
    \url{https://en.wikipedia.org/wiki/Autobiographical_memory} says:
    Autobiographical memory (AM) is a memory system consisting of episodes recollected from an individual's life, based on a combination of episodic (personal experiences and specific objects, people and events experienced at particular time and place) and semantic (general knowledge and facts about the world) memory. It is thus a type of explicit memory.
    
    The autobiographical knowledge base contains knowledge of the self, used to provide information on what the self is, what the self was, and what the self can be. This information is categorized into three broad areas:
    \begin{itemize}
        \item lifetime periods
        \item general events
        \item event-specific knowledge
    \end{itemize}
    
    \section{Types}
        \begin{itemize}
            \item \textbf{Biographical or Personal:}
            These autobiographical memories contain personal biographical information, such as where a person was born or the names of their parents.
        
            \item \textbf{Copies vs.\ Reconstructions}
            \begin{itemize}
                \item \textbf{Copies:} Vivid autobiographical memories of an experience with a considerable amount of visual and sensory-perceptual detail. These memories may vary in authenticity.
                \item \textbf{Reconstructions:} Autobiographical memories that are not direct reflections of raw experiences, but are rebuilt by incorporating new information or interpretations formed in hindsight.
            \end{itemize}
        
            \item \textbf{Specific vs.\ Generic}
            \begin{itemize}
                \item \textbf{Specific:} Memories that contain detailed recollection of a particular event (event-specific knowledge).
                \item \textbf{Generic:} Memories that are vague and contain little detail beyond the type of event that occurred. Episodic autobiographical memories may also fall into this category when one remembered event represents a series of similar events.
            \end{itemize}
        
            \item \textbf{Field vs.\ Observer}
            \begin{itemize}
                \item \textbf{Field memories:} Memories recalled from the original perspective, experienced from a first-person point of view.
                \item \textbf{Observer memories:} Memories recalled from a perspective outside oneself, as if observing the event from a third-person viewpoint. Older memories are often recalled in this form and are more frequently reconstructions, while field memories tend to be more vivid and similar to copies.
            \end{itemize}
        
            \item \url{https://en.wikipedia.org/wiki/Autobiographical_memory#Types}
        \end{itemize}
\end{document}